Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-01-07 Manuscript/TEM_draft.tex
Changed correct topo into best tree's topo

In this study, we focus on the effect of missing data on our ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{}. %CITE-check Lloyd and Friedman or birdy stuff
Although branch length estimation is also important (namely for timing extinction and/or speciation events - e.g. \citealt{ronquista2012}), we do not consider branch lengths in this study. This is partially due to difficulties with simulating branch lengths and topology simultaneously,
but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework. To our knowledge, this question has never been formally assessed.

Here we use a simulation approach to assess the effect of missing data on tree topologies inferred from Total Evidence matrices. Since the molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}, the effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}. Therefore, we focus only on missing data in the morphological part of the matrix. We investigate three major parameters that directly affect the completeness of the morphological part of the matrix: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix. We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods. We chose these parameters because they reflect empirical biases in data availability. The advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis. Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}. We find that the ability of Total Evidence methods to recover the correct topology is increased when using Bayesian methods. Also minimizing the proportion of living taxa with no morphological data and the proportion of missing morphological characters improves the ability of Total Evidence methods to recover the correct topology more so than minimizing the proportion of missing data in the fossil record.

@@

\subsubsection{Summarizing the comparisons}
Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the correct topology in a Total Evidence framework. We also assessed the effect of our missing data parameters jointly with the effects of different tree inference and uncertainty methods (i.e. Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution).

@@

%Explaining the methods comparisons (ANOVA equivalent=among)
Finally, for assessing the effect of the method on the ability to recover the "correct" topology, we compared the similarity between the methods. We calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g. Maximum Likelihood\textit{versus} Bayesian consensus). We then reported the minimum, the 1st quartile, the median, the mean, the 3rd quartile and the maximum of the probabilities of overlap between the methods (i.e. the Bhattacharyya Coefficients) in table~\ref{Tab_Results-Difference_methods}. When two methods perform similarly in their ability to recover the "correct" tree topology (i.e. they display similar normalised metrics scores for each comparison between the "best" tree and the "missing data" trees), we expect the Bhattacharyya Coefficients to be close to 1. On the opposite, if the two methods differ, we expect the Bhattacharyya Coefficients to be close to 0.

@@

%Part 1
As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the correct topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance). However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). In any combination of parameters or metrics, the Maximum Likelihood bootstraps trees and the Bayesian posterior distribution trees performs in a substantially similarly (Table ~\ref{Tab_Results-Difference_methods}) and worth than the Maximum Likelihood and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} - see discussion). %add discussion link

%Part 2
\subsection{Individual effects of missing data parameters} 
As the amount of missing data increases across all three parameters, the ability to recover the correct topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}). However, the decrease of the Normalised Robinson-Fould distance plateaus at minimum of 0.69 for Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values) and there is no substantial difference in topological recovery for 50\% or 75\% of missing living taxa (upper triangle in Fig.~\ref{Fig_Results-paircomp_within}-A). Therefore, clades seems to be more conserved in the Bayesian consensus tree more than any other method (see discussion).  %add discussion link
On the other hand, the decrease in Normalised Triplets distance is less contrasted with a greater overlap of Normalised Triplets distance with the Maximum Likelihood tree (Fig.~\ref{Fig_Results-permeth_perparam}) - mean Bhattacharrya Coefficient=0.10 for the Normalised Robinson-Foulds distance and 0.61 for the Normalised Triplets distance - see table ~\ref{Tab_Results-Difference_methods}).

%Part 3
\subsection{Effect of missing data for fossil taxa ($M_{F}$)}
As the amount of missing data in the fossil part of the matrix increases, the ability to recover the correct topology decreases as well (Fig.~\ref{Fig_Results-permeth_perparam}). For the Robinson-Fould distance, we observe the same pattern than for $M_{L}$ with a better topological recovery of the Bayesian consensus trees on any other methods (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). However, the Normalised Triplets distance drops drastically when going from 25\% of missing data to 50\% (mode) %get the modal values here
making this parameter more sensible to wildcard taxa in the Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} - see discussion). %add discussion link

%Part 4
\subsection{Effect of missing character ($M_{C}$)}
As described for the two previous parameters $M_{L}$ and $M_{F}$, an increase in the amount of missing characters decreases the ability to recover the correct topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}). Similarly again, the Bayesian consensus trees shows higher Normalised Robinson-Fould distance than the Maximum Likelihood tree (Fig.~\ref{Fig_Results-permeth_perparam}). And in a similar way as for the $M_{F}$ parameter, the Normalised Triplet distance drops drastically when going from 25\% of missing data to 50\% (mode). %get the modal values here

%Part 5
\subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
The combine of the missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) reduces the ability to recover the correct topology when each parameters contains a maximum of missing data (i.e. $M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-global_perparam} and see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the combined effects of the parameters in both Normalised Robinson-Foulds and Normalised Triplets distance for the Maximum Likelihood Bootstraps trees and the Bayesian posterior trees distribution). Regarding the clade conservation (i.e. the Normalised Robinson-Foulds distances), the Bayesian consensus tree greatly outperforms the Maximum likelihood trees with Normalised Robinson-Foulds scores always superior than 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the exact values) regardless the combination of missing data parameters (Fig.~\ref{Fig_Results-paircomp_within}-A). However, for the Normalised Triplets distance, the Bayesian consensus tree often displays negative values (i.e. differences greater than expected by chance) but with few substantial differences (Fig.~\ref{Fig_Results-paircomp_within}-B). Also the differences between the Normalised scores of the Maximum Likelihood trees and Bayesian consensus trees are substantially difference for the Normalised Robinson-Foulds distance (mean Bhattacharrya Coefficient=0.10) and for the Normalised Triplets distance (mean Bhattacharrya Coefficient=0.61 - see table ~\ref{Tab_Results-Difference_methods}). Thus the sensitivity of the Bayesian consensus trees to wild card taxa is not highly different than the one for the Maximum likelihood trees (i.e. the distribution of Normalised Triplets distance have a high probability of overlap). The results of the probability of overlap between the pairwise distributions of Normalised Robinson-Foulds and Triplets distances of the Maximum Likelihood trees, the Maximum Likelihood Bootstrap trees and the Bayesian posterior trees distribution is available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.

@@

%As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the correct topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance). 

@@

%The Normalised Robinson-Foulds distance calculate for the trees inferred in Maximum Likelihood framework decreases with the increase of missing data in our three parameters ($M_{L}$, $M_{F}$ or $M_{C}$) with a modal Normalised Robinson-Fould score always below 0.5 when 75\% of the living taxa have no data ($M_{L}$=75\% - Fig.~\ref{Fig_Results-global_perparam}). However, in Bayesian inference, the ability of the method (the Bayesian consensus tree) to recover a "correct" topology plateaus at a minimal Normalised Robinson-Fould score of 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the values distribution). The results of the topological recovery with the parameters combination in both Normalised Robinson-Foulds and Triplets distances for the Maximum Likelihood bootstrap trees and the Bayesian posterior distribution trees are available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.

@@

% NC: Quick comment. You find that Bayesian consensus trees are "best". BUT I think you need to be careful here with what you mean by "best". They are most likely to recover a more similar topology to the best tree than if you do the crazy tree comparison pairs. But that doesn't mean they're best, just that they perform best in your simulations. They are only best for that. If you want to advise people to use a method that best recovers topology as measured using RF or Triplets, then Bayesian consensus is best. BUT generally that is not what people want. Probably need to think some more about this.
% TG: a note about the "correct" topology regarding your comment just after the discussion header: should I replace "correct" by "best" since that's how we're calling the "best" tree anyway (all in quotes)

@@

Our results show that the ability to recover the correct topology in a Total Evidence framework decreases with increasing missing data, regardless of how data were removed or the method of tree inference used. However, these factors affected topological recovery in different ways and to different extents. 
Decreasing the overall number of morphological characters in the matrix ($M_{C}$) and the amount of living taxa with morphological data ($M_{L}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}). Additionally, using Bayesian consensus trees recovered the correct topology more consistently than using Maximum Likelihood trees or Bayesian posterior trees distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
% ----->
% NC: Probably needs another summing up line here. Or maybe not
% <-----
% NC: Make sure it's parameters then tree building methods. NC: Make sure the order of the parameters is always the same L,F, C or whatever! THROUGHOUT the paper. Also correct RF and T to standardised OR normalised RF and T.
% NC: Might want to add a couple of subheadings to match those in the results.
As proposed in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough morphological characters overlaps for most taxa in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}. % NC: What do you mean by data overlap here? That species have overlapping characters?

@@

%Methods
%Part saying the difference between ML and Baycon
When using the Normalised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "correct" topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}). In fact, regardless of the amount of missing data, Bayesian consensus trees always displays the highest Normalised Robinson-Foulds score (always $>$ 0.69; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for details). However, regarding the Normalised Triplets distance, the topological similarity with the "correct" topology decreases in the same way for both the Maximum Likelihood trees and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
%Part saying the difference between BS and baytrees
Regardless of the metric, the Maximum Likelihood boostrap trees and the Bayesian posterior trees distribution always performed poorlier in their ability to recover the "correct" topology, even when no morphological data was missing ($M_{L}$=0\% - Fig.~\ref{Fig_Results-permeth_perparam}). This reflects the fact it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" tree and "best" tree in these cases involved 1000 random pairwise comparisons rather than just one. This will obviously add noise to the results for these two methods.

@@

%Our solutions
Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections. Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "correct" topology. We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections. Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries. However, gaps in the matrix can be filled with future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}. Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "correct" topology (Fig.~\ref{Fig_Results-paircomp_within}). Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved. One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}, so that no one individual collects all the data.

% ----->
% TG: Put that part somewhere in the start with the discussion of ML vs Baycon in the effect of ML part?
%-Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
Variation in our ability to recover the "correct" topology also depended heavily on the method used to infer the tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). Previous studies have shown some superiority of methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}). However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} using Maximum Likelihood and Bayesian methods in a Total Evidence framework. %TG: I guess it's a good selling point: we ARE the first ones to do a sensitivity analysis in a TEM framework (as far as I know).
% <-----

@@

%§7
%Further directions - is that necessary though?
% NC: Cut for now, might want to put these ideas in the appropriate places above instead.
%   -Using real data
%One way to address this issue 
% NC: Which issue???
%would be to run a similar analysis using real data. The amount of studies using Total Evidence data type matrices is steadily increasing \citep[e.g.][]{ronquista2012,slaterphylogenetic2013,beckancient2014} and one could analysis these total evidence matrices in the same framework as this study. However, the true topology (obtained from a matrix with no missing data) is unlikely to be known. A second way to address this issue, and especially to incorporate the variation of signal in morphological character would be to simulate morphological characters at different rates \citep{wrightbayesian2014} or to apply patterns of missing data directly obtained from the fossil record \citep{pattinsonphylogeny2014}. Nevertheless, our simulation already show clear results of the effect of missing data on topology in a Total Evidence framework. Also, our results show how recovering the correct topology can be improved by using Bayesian consensus trees in combination with a maximum amount of characters and a of living taxa with morphological data.
%   -Looking at branch length?

@@

%§Conclusions
% part 1- Fossils are important and can be used combined with living taxa (Intro-§1 & §2)
%   -TEM works with missing data
% part 2- This is how we can improve the data to use them in the best way possible (Conclusion-§1 & §8)
%   -Use bayesian consensus
%   -Go code living taxa
Using Total Evidence matrices with missing data is not a problem for recovering the correct topology as long as enough morphological data is overlapping between the fossil and living taxa. %NC: Explain more clearly what you mean by overlap - of characters?
Topological recovery is still best when there is less missing data, therefore we advise filling as many gaps in Total Evidence matrices as possible. Because this is difficult, if not impossible, for fossil data, we recommend coding as many

>>>>

In this study, we focus on the effect of missing data on our ability to recover a "best" tree's topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{}. %CITE-check Lloyd and Friedman or birdy stuff
Although branch length estimation is also important (namely for timing extinction and/or speciation events - e.g. \citealt{ronquista2012}), we do not consider branch lengths in this study. This is partially due to difficulties with simulating branch lengths and topology simultaneously,
but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework. To our knowledge, this question has never been formally assessed.

Here we use a simulation approach to assess the effect of missing data on tree topologies inferred from Total Evidence matrices. Since the molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}, the effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}. Therefore, we focus only on missing data in the morphological part of the matrix. We investigate three major parameters that directly affect the completeness of the morphological part of the matrix: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix. We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods. We chose these parameters because they reflect empirical biases in data availability. The advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis. Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}. We find that the ability of Total Evidence methods to recover the "best" tree's topology is increased when using Bayesian methods. Also minimizing the proportion of living taxa with no morphological data and the proportion of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree's topology more so than minimizing the proportion of missing data in the fossil record.

@@

\subsubsection{Summarizing the comparisons}
Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the "best" tree's topology in a Total Evidence framework. We also assessed the effect of our missing data parameters jointly with the effects of different tree inference and uncertainty methods (i.e. Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution).

@@

%Explaining the methods comparisons (ANOVA equivalent=among)
Finally, for assessing the effect of the method on the ability to recover the "best" tree's topology, we compared the similarity between the methods. We calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g. Maximum Likelihood\textit{versus} Bayesian consensus). We then reported the minimum, the 1st quartile, the median, the mean, the 3rd quartile and the maximum of the probabilities of overlap between the methods (i.e. the Bhattacharyya Coefficients) in table~\ref{Tab_Results-Difference_methods}. When two methods perform similarly in their ability to recover the "correct" tree topology (i.e. they display similar normalised metrics scores for each comparison between the "best" tree and the "missing data" trees), we expect the Bhattacharyya Coefficients to be close to 1. On the opposite, if the two methods differ, we expect the Bhattacharyya Coefficients to be close to 0.

@@

%Part 1
As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the "best" tree's topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance). However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). In any combination of parameters or metrics, the Maximum Likelihood bootstraps trees and the Bayesian posterior distribution trees performs in a substantially similarly (Table ~\ref{Tab_Results-Difference_methods}) and worth than the Maximum Likelihood and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} - see discussion). %add discussion link

%Part 2
\subsection{Individual effects of missing data parameters} 
As the amount of missing data increases across all three parameters, the ability to recover the "best" tree's topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}). However, the decrease of the Normalised Robinson-Fould distance plateaus at minimum of 0.69 for Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values) and there is no substantial difference in topological recovery for 50\% or 75\% of missing living taxa (upper triangle in Fig.~\ref{Fig_Results-paircomp_within}-A). Therefore, clades seems to be more conserved in the Bayesian consensus tree more than any other method (see discussion).  %add discussion link
On the other hand, the decrease in Normalised Triplets distance is less contrasted with a greater overlap of Normalised Triplets distance with the Maximum Likelihood tree (Fig.~\ref{Fig_Results-permeth_perparam}) - mean Bhattacharrya Coefficient=0.10 for the Normalised Robinson-Foulds distance and 0.61 for the Normalised Triplets distance - see table ~\ref{Tab_Results-Difference_methods}).

%Part 3
\subsection{Effect of missing data for fossil taxa ($M_{F}$)}
As the amount of missing data in the fossil part of the matrix increases, the ability to recover the "best" tree's topology decreases as well (Fig.~\ref{Fig_Results-permeth_perparam}). For the Robinson-Fould distance, we observe the same pattern than for $M_{L}$ with a better topological recovery of the Bayesian consensus trees on any other methods (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). However, the Normalised Triplets distance drops drastically when going from 25\% of missing data to 50\% (mode) %get the modal values here
making this parameter more sensible to wildcard taxa in the Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} - see discussion). %add discussion link

%Part 4
\subsection{Effect of missing character ($M_{C}$)}
As described for the two previous parameters $M_{L}$ and $M_{F}$, an increase in the amount of missing characters decreases the ability to recover the "best" tree's topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}). Similarly again, the Bayesian consensus trees shows higher Normalised Robinson-Fould distance than the Maximum Likelihood tree (Fig.~\ref{Fig_Results-permeth_perparam}). And in a similar way as for the $M_{F}$ parameter, the Normalised Triplet distance drops drastically when going from 25\% of missing data to 50\% (mode). %get the modal values here

%Part 5
\subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
The combine of the missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) reduces the ability to recover the "best" tree's topology when each parameters contains a maximum of missing data (i.e. $M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-global_perparam} and see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the combined effects of the parameters in both Normalised Robinson-Foulds and Normalised Triplets distance for the Maximum Likelihood Bootstraps trees and the Bayesian posterior trees distribution). Regarding the clade conservation (i.e. the Normalised Robinson-Foulds distances), the Bayesian consensus tree greatly outperforms the Maximum likelihood trees with Normalised Robinson-Foulds scores always superior than 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the exact values) regardless the combination of missing data parameters (Fig.~\ref{Fig_Results-paircomp_within}-A). However, for the Normalised Triplets distance, the Bayesian consensus tree often displays negative values (i.e. differences greater than expected by chance) but with few substantial differences (Fig.~\ref{Fig_Results-paircomp_within}-B). Also the differences between the Normalised scores of the Maximum Likelihood trees and Bayesian consensus trees are substantially difference for the Normalised Robinson-Foulds distance (mean Bhattacharrya Coefficient=0.10) and for the Normalised Triplets distance (mean Bhattacharrya Coefficient=0.61 - see table ~\ref{Tab_Results-Difference_methods}). Thus the sensitivity of the Bayesian consensus trees to wild card taxa is not highly different than the one for the Maximum likelihood trees (i.e. the distribution of Normalised Triplets distance have a high probability of overlap). The results of the probability of overlap between the pairwise distributions of Normalised Robinson-Foulds and Triplets distances of the Maximum Likelihood trees, the Maximum Likelihood Bootstrap trees and the Bayesian posterior trees distribution is available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.

@@

%As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the "best" tree's topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance). 

@@

%The Normalised Robinson-Foulds distance calculate for the trees inferred in Maximum Likelihood framework decreases with the increase of missing data in our three parameters ($M_{L}$, $M_{F}$ or $M_{C}$) with a modal Normalised Robinson-Fould score always below 0.5 when 75\% of the living taxa have no data ($M_{L}$=75\% - Fig.~\ref{Fig_Results-global_perparam}). However, in Bayesian inference, the ability of the method (the Bayesian consensus tree) to recover a "best" tree's topology plateaus at a minimal Normalised Robinson-Fould score of 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the values distribution). The results of the topological recovery with the parameters combination in both Normalised Robinson-Foulds and Triplets distances for the Maximum Likelihood bootstrap trees and the Bayesian posterior distribution trees are available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.

@@

% NC: Quick comment. You find that Bayesian consensus trees are "best". BUT I think you need to be careful here with what you mean by "best". They are most likely to recover a more similar topology to the best tree than if you do the crazy tree comparison pairs. But that doesn't mean they're best, just that they perform best in your simulations. They are only best for that. If you want to advise people to use a method that best recovers topology as measured using RF or Triplets, then Bayesian consensus is best. BUT generally that is not what people want. Probably need to think some more about this.
% TG: a note about the "best" tree's topology regarding your comment just after the discussion header: should I replace "correct" by "best" since that's how we're calling the "best" tree anyway (all in quotes)

@@

Our results show that the ability to recover the "best" tree's topology in a Total Evidence framework decreases with increasing missing data, regardless of how data were removed or the method of tree inference used. However, these factors affected topological recovery in different ways and to different extents. 
Decreasing the overall number of morphological characters in the matrix ($M_{C}$) and the amount of living taxa with morphological data ($M_{L}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}). Additionally, using Bayesian consensus trees recovered the "best" tree's topology more consistently than using Maximum Likelihood trees or Bayesian posterior trees distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
% ----->
% NC: Probably needs another summing up line here. Or maybe not
% <-----
% NC: Make sure it's parameters then tree building methods. NC: Make sure the order of the parameters is always the same L,F, C or whatever! THROUGHOUT the paper. Also correct RF and T to standardised OR normalised RF and T.
% NC: Might want to add a couple of subheadings to match those in the results.
As proposed in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough morphological characters overlaps for most taxa in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}. % NC: What do you mean by data overlap here? That species have overlapping characters?

@@

%Methods
%Part saying the difference between ML and Baycon
When using the Normalised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "best" tree's topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}). In fact, regardless of the amount of missing data, Bayesian consensus trees always displays the highest Normalised Robinson-Foulds score (always $>$ 0.69; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for details). However, regarding the Normalised Triplets distance, the topological similarity with the "best" tree's topology decreases in the same way for both the Maximum Likelihood trees and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
%Part saying the difference between BS and baytrees
Regardless of the metric, the Maximum Likelihood boostrap trees and the Bayesian posterior trees distribution always performed poorlier in their ability to recover the "best" tree's topology, even when no morphological data was missing ($M_{L}$=0\% - Fig.~\ref{Fig_Results-permeth_perparam}). This reflects the fact it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" tree and "best" tree in these cases involved 1000 random pairwise comparisons rather than just one. This will obviously add noise to the results for these two methods.

@@

%Our solutions
Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections. Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "best" tree's topology. We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections. Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries. However, gaps in the matrix can be filled with future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}. Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "best" tree's topology (Fig.~\ref{Fig_Results-paircomp_within}). Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved. One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}, so that no one individual collects all the data.

% ----->
% TG: Put that part somewhere in the start with the discussion of ML vs Baycon in the effect of ML part?
%-Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
Variation in our ability to recover the "best" tree's topology also depended heavily on the method used to infer the tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). Previous studies have shown some superiority of methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}). However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} using Maximum Likelihood and Bayesian methods in a Total Evidence framework. %TG: I guess it's a good selling point: we ARE the first ones to do a sensitivity analysis in a TEM framework (as far as I know).
% <-----

@@

%§7
%Further directions - is that necessary though?
% NC: Cut for now, might want to put these ideas in the appropriate places above instead.
%   -Using real data
%One way to address this issue 
% NC: Which issue???
%would be to run a similar analysis using real data. The amount of studies using Total Evidence data type matrices is steadily increasing \citep[e.g.][]{ronquista2012,slaterphylogenetic2013,beckancient2014} and one could analysis these total evidence matrices in the same framework as this study. However, the true topology (obtained from a matrix with no missing data) is unlikely to be known. A second way to address this issue, and especially to incorporate the variation of signal in morphological character would be to simulate morphological characters at different rates \citep{wrightbayesian2014} or to apply patterns of missing data directly obtained from the fossil record \citep{pattinsonphylogeny2014}. Nevertheless, our simulation already show clear results of the effect of missing data on topology in a Total Evidence framework. Also, our results show how recovering the "best" tree's topology can be improved by using Bayesian consensus trees in combination with a maximum amount of characters and a of living taxa with morphological data.
%   -Looking at branch length?

@@

%§Conclusions
% part 1- Fossils are important and can be used combined with living taxa (Intro-§1 & §2)
%   -TEM works with missing data
% part 2- This is how we can improve the data to use them in the best way possible (Conclusion-§1 & §8)
%   -Use bayesian consensus
%   -Go code living taxa
Using Total Evidence matrices with missing data is not a problem for recovering the "best" tree's topology as long as enough morphological data is overlapping between the fossil and living taxa. %NC: Explain more clearly what you mean by overlap - of characters?
Topological recovery is still best when there is less missing data, therefore we advise filling as many gaps in Total Evidence matrices as possible. Because this is difficult, if not impossible, for fossil data, we recommend coding as many
